### 2023-06-19 paper/ATE/cover_letter.tex
ATE update1

\documentclass[a4paper,10pt]{letter}
\usepackage{amsmath}
\usepackage{xcolor}
\linespread{1.2}
\begin{document}
\pagestyle{headings}
\begin{letter}{Dr. Xue-Chao Wang, Managing Guest Editor for Virtual Special Issue, ENERGY}

\opening{Dear Dr. Xue-Chao Wang,}

Enclosed you will find an electronic submission of a manuscript entitled ``Self-tunable approximated explicit MPC: Heat exchanger implementation and analysis''. 
This is an original paper which has neither previously nor simultaneously been submitted to any other journal. This paper is an invited submission to the Virtual Special Issue of ENERGY with title ``Sustainable Use of Energy Contributes to Carbon Emissions Neutrality and Environmental Footprints Reduction''.

We would like to highlight the following contributions of the paper:
\begin{itemize}
	\item Self-tunable control technique does not need any intervention during control
	\item Tuning based on operating conditions compensates for the nonlinear plant behavior
	\item Implementation of the proposed tuning method on a laboratory heat exchanger 
	\item Control performance improved compared to utilizing only one explicit MPC
\end{itemize}

We believe that the presented study is, due to its character, of both industrial and academic interest and within the scope of the journal. We would very much appreciate if you would consider the manuscript for publication in the journal ENERGY.

Sincerely yours, \\
Lenka Gal\v{c}\'{i}kov\'{a}, Petronela Belkov\'{a}, and Juraj~Oravec
% \closing{Sincerely yours,\\Juraj~Oravec, Michaela Horv\'{a}thov\'{a}, and Monika Bako\v{s}ov\'{a}}

\end{letter}
\end{document}