### 2023-07-19 paper/JoPC/cover_letter.tex
cover letter

\documentclass[a4paper,10pt]{letter}
\usepackage{amsmath}
\usepackage{xcolor}
\linespread{1.2}
\begin{document}
\pagestyle{headings}
\begin{letter}{Professor Martin Guay, Editor-in-Chief, Journal of Process Control}

\opening{Dear Professor Martin Guay,}

Enclosed you will find an electronic submission of a manuscript entitled ``Self-tunable approximated explicit MPC: Heat exchanger implementation and analysis''. It is an original paper that has already been submitted to the journal Energy and rejected due to insufficient communication of the proposed method in the results section. We carefully revisited and updated the manuscript based on the reviewers' comments and suggestions, and hope that we managed to make the manuscript more readable and of higher quality.

We would like to highlight the following contributions of the paper:
\begin{itemize}
	\item Self-tunable control technique does not need any intervention during control
	\item Tuning based on operating conditions compensates for the nonlinear plant behavior
	\item Implementation of the proposed tuning method on a laboratory heat exchanger 
	\item Control performance improved compared to utilizing explicit MPC with constant setup
\end{itemize}

We believe that the presented study is, due to its character, of both industrial and academic interest and within the scope of the journal. We would very much appreciate if you would consider the manuscript for publication in the Journal of Process Control.

Sincerely Yours, \\
Lenka Gal\v{c}\'{i}kov\'{a}, Petronela Belkov\'{a}, and Juraj~Oravec
% \closing{Sincerely yours,\\Juraj~Oravec, Michaela Horv\'{a}thov\'{a}, and Monika Bako\v{s}ov\'{a}}

\end{letter}
\end{document}